Introduction to Structural Causal Models

### 2021-05-28 scm_short_presentation.Rmd
scm short presentation

---
title: "Introduction to Structural Causal Models"
author: "Daniel Saggau"
date: "May 2021"
citation_package: --biblatex
bibliography: [ref.bib]
#csl: ieee.csl
output: 
  beamer_presentation:
    theme: Madrid
header-includes:
- \definecolor{UBCblue}{rgb}{0.04706, 0.13725, 0.26667}
- \usecolortheme[named=UBCblue]{structure}
---

@@

# Example

```{tikz,fig.cap="Structural Causal Model", fig.align="center", echo =F}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
  \begin{tikzpicture}[
    sharp corners=2pt,
    inner sep=6pt,
    node distance=2cm,
    >=latex]
\tikzstyle{my node}=[draw, shape = circle, minimum height=1cm,minimum width=1cm]
\tikzstyle{latent}=[draw, shape = rectangle, minimum height=1cm,minimum width=1cm]
\node[my node, fill=gray!30] (A){A};
\node[latent,left of=A](UA){$U_A$};
\node[my node, fill=gray!30] at ($(A)!0.5!(A)-(0pt,1.5cm)$) (E) {E};
\node[my node,right = 1 cm of E, fill=gray!30](C){C};
\node[latent,left = 1cm of E](UE){$U_E$};
\node[latent,right = 1cm of A](UC){$U_C$};
\draw[->] (A) -- (E);
\draw[->] (UA) -- (A);
\draw[->] (UE) -- (E);
\draw[->] (UC) -- (C);
\draw[->] (A) -- (C);
\draw[->] (E) -- (C);
\end{tikzpicture}
```

@@

#

>>>>

---
title: "Introduction to Structural Causal Models"
author: "Daniel Saggau"
date: "May 2021"
citation_package: --biblatex
output: 
  beamer_presentation:
    theme: Madrid
header-includes:
- \definecolor{UBCblue}{rgb}{0.04706, 0.13725, 0.26667}
- \usecolortheme[named=UBCblue]{structure}
---

@@

# Assumptions SCM

- Noise terms independent 
- Mechanisms independent/autonomous 
- Set of equations 
- Endogenous and Exogenous Variables

# Graphical Example

\begin{columns}
\begin{column}{0.48\textwidth}

```{tikz,fig.cap="Structural Causal Model", fig.align="center", echo =F}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
  \begin{tikzpicture}[
    sharp corners=2pt,
    inner sep=6pt,
    node distance=2cm,
    >=latex]
\tikzstyle{my node}=[draw, shape = circle, minimum height=1cm,minimum width=1cm]
\tikzstyle{latent}=[draw, shape = rectangle, minimum height=1cm,minimum width=1cm]
\node[my node, fill=gray!30] (N){N};
\node[latent,left of=N](UN){$U_N$};
\node[my node, fill=gray!30] at ($(N)!0.5!(N)-(0pt,1.5cm)$) (L) {L};
\node[my node,right = 1 cm of L, fill=gray!30](E){E};
\node[latent,left = 1cm of L](UL){$U_L$};
\node[latent,right = 1cm of N](UE){$U_E$};
\draw[->] (UN) -- (N);
\draw[->] (UL) -- (L);
\draw[->] (UE) -- (E);
\draw[->] (N) -- (E);
\draw[->] (L) -- (E);
\end{tikzpicture}
```

\end{column}
\begin{column}{0.48\textwidth}
$\{U_N\}$ - Numeric Ability

$\{U_L\}$ - Literary Ability

$\{U_C\}$ - Quality of the Student

$\{A\}$ - Math Test

$\{E\}$ - English-Test

$\{C\}$ - GPA

\end{column}
\end{columns}

# Mathematical Example

\begin{columns}
\begin{column}{0.48\textwidth}

$N:= f_N(U_N)$\newline
$L:=f_L(U_L)$\newline
$E:= f_E(N,L,U_E)$\newline

\end{column}
\begin{column}{0.48\textwidth}

$\{U_N\}$ - Numeric Ability

$\{U_L\}$ - Literary Ability

$\{U_E\}$ - Quality of the Student

$\{N\}$ - Math Test

$\{L\}$ - English-Test

$\{E\}$ - GPA

\end{column}
\end{columns}

@@

# Questions for Discussion
